---
title: "Effects on the Fuel Efficiency of Popular Car Brands"
format: docx
editor: visual
---

## MPG Data

```{r}
library(ggplot2)
data(mpg)
mpg
model <- lm(hwy ~ displ + cyl + drv + year, data = mpg)

```

## Model Diagnostics

## Residual Diagnostics/Residuals vs Fitted (homoscedasticity)

```{r}
plot(model, which = 1)
```

There is no pattern so homoscedasticity is satisfied.

## Predicted Values and Residuals

```{r}
predictions <- predict(model)
residuals <- model$residuals

```

## Normality of Residuals

```{r}
qqnorm(residuals)
qqline(residuals)
```

For the most part, the points stay on the line so the residuals are normally distributed.

## Independence of Errors

```{r}
plot(residuals, main = "Residuals vs Index", xlab = "Index", ylab = "Residuals")
abline(h = 0, col = "red")
```

There is no trend so independence of errors is satisfied.

## Multicollinearity

```{r}
library(car)
vif(model)
```

The predictors show low to moderate multicollinearity.

## Outliers

```{r}
std_residuals_model <- rstandard(model)  
plot(std_residuals_model, main = "Standardized Residuals", xlab = "Index", ylab = "Standardized Residuals")
abline(h = c(-2, 2), col = "red")
```

> Based on the standardized residual plot, a few observations fall outside the ±2 range, indicating potential outliers. There are only three points that are extreme residuals beyond +3 but it is a low number so they are unlikely to have a substantial influence on the regression model.

## Leverage and Influential Points

```{r}
leverage <- hatvalues(model)
plot(leverage, main = "Leverage Values", xlab = "Index", ylab = "Leverage")
abline(h = 2 * mean(leverage), col = "red")
```

Most points fall below the red line which means the points are not overly influential on their own. Points above the red line have high leverage.

## Scatter plots to explore relationship between Engine Displacement and Hwy

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Engine Displacement vs Highway MPG", x = "Engine Displacement (displ)", y = "Highway MPG (hwy)")
```

As engine displacement increases, the highway MPG decreases which shows a negative linear relationship.

## Scatter plots for Cyl and Hwy

```{r}
ggplot(mpg, aes(x = cyl, y = hwy)) + 
  geom_boxplot() +
  labs(title = "Number of Cylinders vs Highway MPG", x = "Cylinders (cyl)", y = "Highway MPG (hwy)")

```

Vehicles with more cylinders generally have lower highway MPG.

```{r}
summary(model)

confint(model)
```
